---
title: "Assignment_1_rxh"
author: "Rongxin Hu  2219247"
date: "2025-03-06"
documentclass: ctexart
keywords:
  - 中文
  - R Markdown
output:
  rticles::ctex:
    fig_caption: yes
    number_sections: yes
    toc: yes
---


# Load Necessary Packages

```{r}
library(httr)
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(data.table)

```


# The FRED-MD Dataset & The Transformations

## Load FRED-MD Dataset

```{r }
url <- "https://www.stlouisfed.org/-/media/project/frbstl/stlouisfed/research/fred-md/monthly/current.csv?sc_lang=en&hash=80445D12401C59CF716410F3F7863B64"

response <- GET(url)
if (http_error(response)) {
  stop("Failed to download file.")
} else {
  df <- read_csv(content(response, as = "text", encoding = "UTF-8"))
}
print(df)

#This method bypasses the HTTP connection issues that come with R
```


## Extract Transformation Codes

```{r }
transformation_codes <- data.frame(Series = names(df)[-1], Transformation_Code = as.numeric(df[1, -1]))

```


## Function to Apply Transformations 

```{r }
mdiff <- function(x) {
    x - dplyr::lag(x, 1, default = NA)
}


apply_transformation <- function(series, code) {
  if (code == 1) {
    return(series)
  } else if (code == 2) {
    return(mdiff(series))
  } else if (code == 3) {
    return(mdiff(mdiff(series)))
  } else if (code == 4) {
    return(log(series))
  } else if (code == 5) {
    return(mdiff(log(series)))
  } else if (code == 6) {
    return(mdiff(mdiff(log(series))))
  } else if (code == 7) {
    return(mdiff(series) / dplyr::lag(series, 1) - 1)
  } else {
    stop("Invalid transformation code")
  }
}

```


## Applying the Transformations to Each Column 

```{r }
for (i in 1:nrow(transformation_codes)) {
  series_name <- transformation_codes$Series[i]
  code <- transformation_codes$Transformation_Code[i]
  df[[series_name]] <- apply_transformation(as.numeric(df[[series_name]]), code)
}

df_cleaned <- df[-c(1:3), ]

```


## Plot Transformed Series

```{r }
series_to_plot <- c('INDPRO', 'CPIAUCSL', 'TB3MS')
series_names <- c('Industrial Production', 'Inflation (CPI)', '3-month Treasury Bill rate')

 plot_data <- df_cleaned %>%
   select(sasdate, all_of(series_to_plot)) %>%
   pivot_longer(-sasdate, names_to = "series", values_to = "value") %>%
   mutate(sasdate = mdy(sasdate),
          series_name = factor(series, levels = series_to_plot, labels = series_names))

 ggplot(plot_data, aes(x = sasdate, y = value, color = series_name)) +
   geom_line() +
   facet_wrap(~series_name, scales = "free", ncol=1) +
   theme_minimal() +
   labs(x = "Year", y = "Transformed Value") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")


```


# Forecasting in Time Series

## Prepare Data for Estimation

```{r }
Yraw <- df_cleaned$INDPRO
Xraw <- df_cleaned %>% select(CPIAUCSL, TB3MS)

num_lags <- 4  # this is p
num_leads <- 1  # this is h

X <- data.frame(Ones = rep(1, nrow(df_cleaned)))

for (lag in 0:num_lags) {
  X[paste0('INDPRO_lag', lag)] <- dplyr::lag(Yraw, lag)
}

for (col in names(Xraw)) {
  for (lag in 0:num_lags) {
    X[paste0(col, '_lag', lag)] <- dplyr::lag(Xraw[[col]], lag)
  }
}

y <- dplyr::lead(Yraw, num_leads)

```

## Getting the Last Row fro Forecasting

```{r }
X_T <- as.matrix(tail(X, 1))
```


## Removing NA Rows (Due to Lagging/Leading)

```{r }
complete_cases <- complete.cases(X, y)
X <- X[complete_cases, ]
y <- y[complete_cases]
```


## Estimation and Forecast

```{r }
y <- as.vector(y)
X <- as.matrix(X)

beta_ols <- solve(crossprod(X), crossprod(X, y))
```


## Produce the One Step Ahead Forecast

```{r }
forecast <- (X_T %*% beta_ols) * 100

```

# My Forecasting Exercise 

## Defining Forecasting Functions (from hint)

```{r }

calculate_msfe <- function(y, X, num_lags, h) {
  X <- as.matrix(X)  
  n <- length(y)
  errors <- numeric(n - num_lags - h)
  
  for (t in (num_lags + 1):(n - h)) {
    # construct lag matrix
    y_lags <- embed(y[1:t], num_lags + 1)[, -1, drop = FALSE]
    X_lags <- lapply(1:ncol(X), function(i) {
      x_col <- X[1:t, i]
      embed(x_col, num_lags + 1)[, -1, drop = FALSE]
    })
    
    # combined lagged variables
    X_lags <- do.call(cbind, X_lags)
    
    #  # design mayrix (including intercept term)
    design_matrix <- cbind(1, y_lags, X_lags)
    
    # fix the index range of y_target
    y_target <- y[(num_lags + 1 + h):(t + h)]  
    
    complete_cases <- complete.cases(design_matrix, y_target)
    design_matrix <- design_matrix[complete_cases, , drop = FALSE]
    y_target <- y_target[complete_cases]
    
    # ensure sufficient number of rows in the design matrix
    if (nrow(design_matrix) > ncol(design_matrix)) {
      beta <- solve(t(design_matrix) %*% design_matrix) %*% t(design_matrix) %*% y_target
      X_T <- c(1, tail(y_lags, 1), tail(X_lags, 1))
      forecast <- X_T %*% beta
      errors[t - num_lags] <- (y[t + h] - forecast)^2
    } else {
      errors[t - num_lags] <- NA
    }
  }
  
  MSFE <- mean(errors, na.rm = TRUE)
  return(MSFE)
}

```


## Different h to Our Model

```{r }
msfe_h1 <- calculate_msfe(Yraw, Xraw, num_lags = 4, h = 1)
msfe_h4 <- calculate_msfe(Yraw, Xraw, num_lags = 4, h = 4)
msfe_h8 <- calculate_msfe(Yraw, Xraw, num_lags = 4, h = 8)

print(msfe_h1)
print(msfe_h4)
print(msfe_h8)

```


## Different p to Our Model

```{r }
msfe_p1 <- calculate_msfe(Yraw, Xraw, num_lags = 1, h = 4)
msfe_p4 <- calculate_msfe(Yraw, Xraw, num_lags = 4, h = 4)
msfe_p8 <- calculate_msfe(Yraw, Xraw, num_lags = 8, h = 4)

print(msfe_p1)
print(msfe_p4)
print(msfe_p8)


```

